### 2025-01-07 images/encoders/quantikz/IQP.tex
manuscript

\documentclass[preview, border=1mm]{standalone}
\usepackage{tikz} %Format quantum circuits
\usetikzlibrary{quantikz2}%Format quantum circuits
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools,physics}
\def\myvdots{\ \vdots\ }

\begin{document}
\begin{quantikz} 
\lstick{$q_0$} & \gate{H} & \gate{R_{Z} (x_{0})} & \gate[2,disable auto height]{ZZ(x_{0},x_{1})} &  & \gate[3,disable auto height]{ZZ(x_{0},x_{2})} & \gate[5,disable auto height]{ZZ(x_{0},x_{n})}   & & &  \\ 
\lstick{$q_1$} & \gate{H} & \gate{R_{Z} (x_{1})} & & \gate[2,disable auto height]{ZZ(x_{1},x_{2})}  & & &  \gate[4,disable auto height]{ZZ(x_{1},x_{n})}  &  &\\ 
\lstick{$q_2$} & \gate{H} & \gate{R_{Z} (x_{2})} & & & & & & \gate[3,disable auto height]{ZZ(x_{2},x_{n})}  & \\ 
\setwiretype{n} \lstick{}  & \myvdots & \myvdots & \myvdots  & \myvdots & \myvdots & \\
\lstick{$q_n$}& \gate{H}  & \gate{R_{Z} (x_{n})} & & & & & & &\\
\end{quantikz}
\end{document}