### 2025-01-08 images/ansatz/quantikz/Full-CRX.tex
manuscript images

\documentclass[preview, border=1mm]{standalone}
\usepackage{tikz} %Format quantum circuits
\usetikzlibrary{quantikz2}%Format quantum circuits
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools,physics}
\def\myvdots{\ \vdots\ }

\begin{document}

%Full-CRX
\resizebox{\textwidth}{!}{
\begin{quantikz}
	\lstick{$q_0$}& \gate{RX} & \gate{RZ} &\ctrl{1}& \ctrl{2}&\ctrl{3}&\ctrl{4}&\gate{RX} &&&&\gate{RX}&&&&\gate{RX}& &&&\gate{RX}&&&&\\
	\lstick{$q_1$}& \gate{RX} & \gate{RZ} & \gate{RX} &&&& \ctrl{-1} & \ctrl{1} & \ctrl{2} &  \ctrl{3}&&\gate{RX}&&&&\gate{RX}&&&&\gate{RX}&&&\\
	\lstick{$q_2$}& \gate{RX} & \gate{RZ} && \gate{RX}&&&&\gate{RX}&&& \ctrl{-2}&\ctrl{-1}&\ctrl{1}&\ctrl{2}&&&\gate{RX}&&&&\gate{RX}&&\\
	\lstick{$q_3$}& \gate{RX} & \gate{RZ} &&&\gate{RX}&&&&\gate{RX}&&&&\gate{RX}&&\ctrl{-3}&\ctrl{-2}&\ctrl{-1}& \ctrl{1}&&&&\gate{RX}&\\
	\lstick{$q_4$}& \gate{RX} & \gate{RZ} &&&&\gate{RX}&&&&\gate{RX}&&&&\gate{RX}&&&&\gate{RX}&\ctrl{-4}&\ctrl{-3}&\ctrl{-2}&\ctrl{-1}&\\
\end{quantikz}
}

\end{document}