### 2025-02-10 images/ansatz/quantikz/Full-CRX.tex
Mon 10 Feb 2025 17:00:20 EST poster/manuscript figures

\documentclass[preview, border=1mm]{standalone}
\usepackage{tikz} %Format quantum circuits
\usetikzlibrary{quantikz2}%Format quantum circuits
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools,physics}
\def\myvdots{\ \vdots\ }

\begin{document}

%Full-CRX(	\theta)
\resizebox{\textwidth}{!}{
\begin{quantikz}
	\lstick{$q_0$}& \gate{RX(	\theta)} & \gate{RZ(	\theta)} &\ctrl{1}& \ctrl{2}&\ctrl{3}&\ctrl{4}&\gate{RX(	\theta)} &&&&\gate{RX(	\theta)}&&&&\gate{RX(	\theta)}& &&&\gate{RX(	\theta)}&&&&\\
	\lstick{$q_1$}& \gate{RX(	\theta)} & \gate{RZ(	\theta)} & \gate{RX(	\theta)} &&&& \ctrl{-1} & \ctrl{1} & \ctrl{2} &  \ctrl{3}&&\gate{RX(	\theta)}&&&&\gate{RX(	\theta)}&&&&\gate{RX(	\theta)}&&&\\
	\lstick{$q_2$}& \gate{RX(	\theta)} & \gate{RZ(	\theta)} && \gate{RX(	\theta)}&&&&\gate{RX(	\theta)}&&& \ctrl{-2}&\ctrl{-1}&\ctrl{1}&\ctrl{2}&&&\gate{RX(	\theta)}&&&&\gate{RX(	\theta)}&&\\
	\lstick{$q_3$}& \gate{RX(	\theta)} & \gate{RZ(	\theta)} &&&\gate{RX(	\theta)}&&&&\gate{RX(	\theta)}&&&&\gate{RX(	\theta)}&&\ctrl{-3}&\ctrl{-2}&\ctrl{-1}& \ctrl{1}&&&&\gate{RX(	\theta)}&\\
	\lstick{$q_4$}& \gate{RX(	\theta)} & \gate{RZ(	\theta)} &&&&\gate{RX(	\theta)}&&&&\gate{RX(	\theta)}&&&&\gate{RX(	\theta)}&&&&\gate{RX(	\theta)}&\ctrl{-4}&\ctrl{-3}&\ctrl{-2}&\ctrl{-1}&\\
\end{quantikz}
}

\end{document}