### 2025-02-10 images/encoders/quantikz/IQP.tex

\documentclass[preview, border=1mm]{standalone}
\usepackage{tikz} %Format quantum circuits
\usetikzlibrary{quantikz2}%Format quantum circuits
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools,physics}

\begin{document}
	\begin{quantikz} 
	\lstick{$q_0$} & \gate{H} & \gate{R_{Z} (x_{0})} & \gate[2,disable auto height]{ZZ(x_{0},x_{1})} &   \gate[3,disable auto height]{ZZ(x_{0},x_{2})} & \gate[4,disable auto height]{ZZ(x_{0},x_{3})} &  \gate[5,disable auto height]{ZZ(x_{0},x_{4})}   & &  & & & & & \\ 
	\lstick{$q_1$}& \gate{H}  & \gate{R_{Z} (x_{1})} & & & & & \gate[2,disable auto height]{ZZ(x_{1},x_{2})} & \gate[3,disable auto height]{ZZ(x_{1},x_{3})} & \gate[4,disable auto height]{ZZ(x_{1},x_{4})} & & & &\\
	\lstick{$q_2$}& \gate{H}  & \gate{R_{Z} (x_{2})} & & & & & & & &  \gate[2,disable auto height]{ZZ(x_{2},x_{3})} &  \gate[3,disable auto height]{ZZ(x_{2},x_{4})} & & \\
	\lstick{$q_3$}& \gate{H}  & \gate{R_{Z} (x_{3})} & & & & & & & & & & \gate[2,disable auto height]{ZZ(x_{3},x_{4})} & \\
	\lstick{$q_4$}& \gate{H}  & \gate{R_{Z} (x_{4})} & & & & & & & & & & & \\
	\end{quantikz}
\end{document}